---
title: "Tables"
author: "Rodney Murray"
format: 
  html:
    toc: true
editor: visual
---

```{r, warning=F, message=F}
library(dplyr)
library(ggplot2)
library(plotly)
library(viridis)
library(gridExtra)
library(tidyr)
library(knitr)
library(kableExtra)
library(sparkline)
library(DT)
library(reactable)
library(stringr)
```


## Description

I downloaded corn yield data from [USDA's quick stats lite site](https://www.nass.usda.gov/Quick_Stats/Lite/index.php#91CC0BA8-1B17-3774-91BB-3549B470CD60) so that I could compare corn yields for irrigated land over time from a couple of states of interest: Montana (where I will be moving to) and Colorado (because of CSU).

To replicate the data one can visit the site linked above and apply the filters described below:

-   Sector = CROPS
-   Group = FIELD CROPS
-   Commodity = CORN
-   View = Acreage, Yield, and Production - Irrigated / Non-Irrigated
-   Year = 1950-2022
-   Geographic Level = State

**Note:** I exported data for all states, but had to do it in two exports. The tool provided an empty CSV when I attempted to select data for all states at once. This data is also available in this GitHub repo.

## Load Data

```{r}
# Read data from a csv file
# data <- read.csv('corn_production.csv')
data1 <- read.csv('corn_production1.csv')
data2 <- read.csv('corn_production2.csv')
data <- bind_rows(data1, data2)

# Change the names to lower case so that they're easier for me to work with
names(data) <- tolower(names(data))
```

## Data preparation

Here I will rename, select a subset of columns, and filter the data to records that have data for production as well as those where the corn was grown on irrigated land. I will also filter data at this point to Year \> 1980 to help focus on the time period I'm interested in.

```{r, warning=FALSE}
# Rename, select a subset of columns, and filter
prep_data <- data %>%
  rename(
    production=production.in.bu, 
    harvested_area=area.harvested.in.acres,
    yield=yield.in.bu...acre
    ) %>% 
  filter(
    production > 0,
    prodn.practice == 'IRRIGATED',
    year >= 1980
    ) %>% 
  select(year, location, prodn.practice, harvested_area, production, yield) %>% 
  mutate(
    harvested_area=as.numeric(str_replace_all(harvested_area, ',', '')),
    production=as.numeric(str_replace_all(production, ',', '')),
    location = str_to_title(location)
  )
```


```{r, warning=FALSE}
# Get the average yields for stats in the 2000 to 2005 time period
average_state_yields <- prep_data %>% filter(
      year >= 2000,
    year < 2005
  ) %>% 
  group_by(location) %>% 
  summarize(
    harvested_area=mean(harvested_area),
    production=mean(production),
    yield=mean(yield)
    ) %>% 
  mutate(state = location)
```


## Table

```{r}
table_stats <- average_state_yields %>% 
  select(state, harvested_area, production, yield)
cols <- c('State', 'Harvested Area', 'Production', 'Yield')
units <- c('', 'acre', 'bu', 'bu/acre')

kable(
  table_stats,
  col.names=units,
  align='c',
  digits=1
  ) %>% 
  add_header_above(cols) %>% 
  kable_material(lightable_options = c('hover', 'striped', 'condensed'))
```

